---
title: "Project 2"
author: "Jamie Dunwoodie-Rowley"
output:
  html_document:
    code_folding: hide
subtitle: "STATS 220 Semester One 2024"
---

```{r setup, include=FALSE}
# Load required libraries
library(tidyverse)

# Set chunk options
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, error=FALSE)
```

## Introduction

## Analytics

Three things i learnt about the data:

-   I learnt that the highest common personal cost is phone plans, this makes sense as often poeple living at home still have to pay for there phone plans

-   I learnt that the lowest personal cost category was travel, I believe this is because people dont travel on a week to week basis, instead they save for a few big trip.

-   very few people save more than 50% of there income a week it is the least saved amount

```{r, include=TRUE}
#load data...
learning_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTkTVJOlnhruvrKBti7-64d16vJkLC9Lcyr5HHe_HbsoBE-cq-EEVgOQnrwsahoVh4PMZmzH5yO95W0/pub?output=csv")

#Rename cols
learning_data <- learning_data %>%
  rename(total_income=2,living_out=3,living_costs=4,personal_expences=5,personal_costs=6,expected_savings=7)
#calculate total income
total_income <- sum(learning_data$total_income, na.rm = TRUE) # had to remove na values as some people didnt put in answers before i made them all required! 

# Calculate average personal spend
avr_personal_spend <- sum(learning_data$personal_expenses, na.rm = TRUE) / sum(!is.na(learning_data$personal_expenses)) #had to google this is.na function!! 
# Calculate average living spend
avr_living_spend <- sum(learning_data$living_out, na.rm = TRUE) / sum(!is.na(learning_data$living_out))

print(paste("The total income of all participants was $", total_income, sep=""))
print(paste("Average personal spend $",avr_personal_spend, ", Average living spend $",avr_living_spend, sep=""))

learning_data %>%
  ggplot() +
  geom_bar(aes(x = total_income, fill=expected_savings))

utilities <- learning_data %>% 
  separate_rows(living_costs, sep = ", ")

utilities %>%
  ggplot() +
  geom_bar(aes(x = living_costs))
```

## Creativity

## Learning reflection
